Defining properties regulating pioneer-factor binding and function

### 2023-01-04 manuscript/manuscript.qmd
update MS results section

| Reagent                               | Source/vendor             | Catalog/accession \# | Notes                   |
|-------------------|------------------|------------------|------------------|
| pMT-puro                              | AddGene                   | 17923                |                         |
| Gibson Cloning                        | New England Biolabs       | E2611                |                         |
| Schneider's Medium                    | Thermo Fisher Scientific  | 21720001             |                         |
| Fetal Bovine Serum                    | Omega Scientific          | FB-02                |                         |
| Antibiotic/antimycotic                | Thermo Fisher Scientific  | 15240062             |                         |
| Effectene Transfection Reagent        | Qiagen                    | 301427               |                         |
| PVDF Membrane                         |                           |                      |                         |
| Formaldehyde                          |                           |                      |                         |
| Protease inhibitor tablets            |                           |                      |                         |
| NEBNext Ultra II DNA library prep kit |                           |                      |                         |
| TruSeq v2 RNA library prep kit        |                           |                      |                         |
| Trizol                                |                           |                      |                         |
| **Antibodies**                        |                           |                      |                         |
| Anti-Zld                              | Harrison et al., 2010     |                      | ChIP: 6 µL , WB: 1:750  |
| Anti-Grh                              | Harrison et al., 2010     |                      | ChIP: 8 µL, WB: 1:1000  |
| Anti-Twi                              | He et al., 2015           |                      | ChIP: 10 µg, WB: 1:1000 |
| Anti-HA-peroxidase clone 3F10         | Roche                     | 12013819001          | WB: 1:500               |
| Anti-Tubulin DM1A                     | Sigma                     | T6199-200UL          | WB: 1:5000              |
| goat anti-rabbit IgG-HRP conjugate    | Bio-Rad                   | 1706515              | WB: 1:3000              |
| Anti-HA clone 12CA5                   | Roche                     |                      | ChIP: 7.5 µL            |
| Anti-H3K27me3                         | Cell Signaling Technology |                      | WB: 1:1000, CR: 0.5 µL  |
|                                       |                           |                      |                         |
|                                       |                           |                      |                         |

## Cloning

@@

-   **TG2A- Zld RB**

-   **TG3A- Grh PH**

-   **TG4A**

-   **TG6A**

-   **TG8A**

-   **MF8A**

-   GFP-DBD lines

@@

Bowtie2 version 2.4.4 [@Langmead2012] was used to align ChIP-seq reads to the *Drosophila melanogaster* genome (version dm6) with the following non-default parameters: -k 2 \--very-sensitive \--no-mixed \--no-discordant -X 5000. Aligned reads were filtered to include only reads with a mapping quality score \> 30. Reads mapping to unplaced scaffolds or the mitochondrial genome were discarded.

peak filtering

generation of bigwigs

normalization of bigwigs

@@

@@

@@

@@

@@

@@

>>>>

| Reagent                               | Source/vendor              | Catalog/accession \# | Notes                   |
|---------------------|-----------------|-----------------|-----------------|
| pMT-puro                              | AddGene                    | 17923                |                         |
| Gibson Cloning                        | New England Biolabs        | E2611                |                         |
| Schneider's Medium                    | Thermo Fisher Scientific   | 21720001             |                         |
| Fetal Bovine Serum                    | Omega Scientific           | FB-02                |                         |
| Antibiotic/antimycotic                | Thermo Fisher Scientific   | 15240062             |                         |
| Effectene Transfection Reagent        | Qiagen                     | 301427               |                         |
| PVDF Membrane                         |                            |                      |                         |
| Formaldehyde                          |                            |                      |                         |
| Protease inhibitor tablets            |                            |                      |                         |
| NEBNext Ultra II DNA library prep kit |                            |                      |                         |
| TruSeq v2 RNA library prep kit        |                            |                      |                         |
| Trizol                                |                            |                      |                         |
| **Antibodies**                        |                            |                      |                         |
| Anti-Zld                              | Harrison et al., 2010      |                      | ChIP: 6 µL , WB: 1:750  |
| Anti-Grh                              | Harrison et al., 2010      |                      | ChIP: 8 µL, WB: 1:1000  |
| Anti-Twi                              | He et al., 2015            |                      | ChIP: 10 µg, WB: 1:1000 |
| Anti-HA-peroxidase clone 3F10         | Roche                      | 12013819001          | WB: 1:500               |
| Anti-Tubulin DM1A                     | Sigma                      | T6199-200UL          | WB: 1:5000              |
| goat anti-rabbit IgG-HRP conjugate    | Bio-Rad                    | 1706515              | WB: 1:3000              |
| Anti-HA clone 12CA5                   | Roche                      |                      | ChIP: 7.5 µL            |
| Anti-H3K27me3                         | Cell Signaling Technology  | C36B11; Lot #19      | WB: 1:1000, CR: 0.5 µL  |
|                                       |                            |                      |                         |
|                                       |                            |                      |                         |

##

@@

@@

Bowtie2 version 2.4.4 [@Langmead2012] was used to align ChIP-seq reads to the *Drosophila melanogaster* genome (version dm6) with the following non-default parameters: -k 2 \--very-sensitive \--no-mixed \--no-discordant -X 5000.
Aligned reads were filtered to include only reads with a mapping quality score \> 30.
Reads mapping to unplaced scaffolds or the mitochondrial genome were discarded.
Peak calling was performed using MACS2 version 2.2 with parameters -g dm \--call-summits.
Only peaks that were detected in all replicates were considered in downstream analysis.
bigWig files were generated using deepTools bamCoverage version 3.5.1 with parameters --binSize 10.
bigWig files were z-score normalized by subtracting the genome-wide average of all bins from each bin value and dividing by the standard deviation of all bins.

@@

ATAC-seq was performed as described previsouly [@Buenrostro2013].
2×10^5^ cells were harvested for ATAC-seq by centrifugation at 600 × *g* for 3 minutes at RT.
Cells were washed once in 100 µL 1X PBS and resuspended in 100 µL ATAC lysis buffer (10 mM Tris pH 7.5, 10 mM NaCl, 3 mM MgCl~2~,0.1% NP-40).
Cells were centrifuged at 600 × *g* for 10 minutes at 4°C.
Supernatant was removed and pellet was resuspended in 47.5 µL buffer TD (Illumina).
2.5 uL of Tagment DNA Enzyme (Illumina) was added and samples were incubated at 37°C for 30 minutes.
Tagmented DNA was purified using MinElute Reaction Cleanup Kit (Qiagen) and eluted in 10 µL of buffer EB.
DNA was PCR amplified for 12 cycles with the following conditions:72°C for 5 minutes, 98°C for 30 seconds, then 12 cycles of 98°C for 10 seconds, 63°C for 30 seconds and 72°C for 1 minute.
Amplified libraries were purified using a 1.2X ratio of Axygen paramagnetic beads.
Sequencing was performed on the Illumina NovaSeq 6000 using 150bp paired-end reads.

@@

Raw ATAC-seq reads were trimmed to remove adapter sequences using NGmerge version 0.3 [@Gaspar2018].
Trimmed reads were aligned to the *Drosophila melanogaster* genome (version dm6) using bowtie2 with paramters -k 2 \--very-sensitive \--no-mixed \--no-discordant -X 5000.
Reads with a mapping quality \< 30 and reads aligning to the mitochondrial genome or scaffolds were discarded.
As described previously, only fragments \< 100bp were considered for downstream analysis [@Buenrostro2013].
Reads were combined across all replicates and peak calling was performed using MACS2 with parameters -f BAMPE \--keep-dup all -g dm \--call-summits.
featureCounts was used to count the number of reads aligning within 200bp (100bp upstream or downstream) of peak summits.
Differential accessibility analysis was performed using DESeq2.
Regions with an adjusted p-value \< 0.05 were considered differentially accessible.

Cells were plated at 1×10^6^ cells per mL and DMSO or tazemetostat was added to a final concentration of 10 µM.
Cells were split on days 2 and 5 and tazemetostat was added to maintain a final concentration of 10 µM.
On day 5, CuSO~4~ was added to achieve physiological expression of Zld, Grh or Twi (see above).
On day 7, cells were harvested for immunoblotting, ChIP and CUT&RUN.

@@

CUT&RUN was performed using EpiCypher reagents according to the manufacturer protocol.
2×10^5^ cells were used for each CUT&RUN reaction.
Overnight incubation with antibodies was performed in antibody buffer containing 0.05% digitonin.
0.5 µL anti-H3K27me3 (Cell signaling technologies) or rabbit IgG were used for CUT&RUN.
Prior to addition of antibody, a 1:5 dilution of K-MetStat Panel spike-in nucleosomes (EpiCypher, cat. \# 19-1002) was added to each reaction.
Libraries were prepared using the NEBNext Ultra II kit (NEB).
During library preparation, cleanup steps were performed using a 1.1X ratio of Axygen paramagnetic beads, as recommended by EpiCypher.
PCR amplification was performed with the following conditions: 98°C for 45 seconds, then 14 cycles of 98°C for 15 seconds, 60°C for 10 seconds, and a final extension of 72°C for 1 minute.
Sequencing was performed on the Illumina NovaSeq 6000 with 150bp paired-end reads.

@@

Trimming

bigwigs

z-score normalization

spike-in normalization

@@

Previously published datasets analyzed in this study can be found in supplemental table XXXX.
All previously published data were analyzed in parallel with data from this study using analysis parameters described above.

@@

Sequencing data have been deposited in the Gene Expresion Omnibus under accession GSEXXXX.
All analysis code is available on GitHub: https://github.com/tjgibson/S2_pioneers_manuscript.
An archived version of the data processing and analysis pipeline is available on Zenodo: doiXXXX.
